---
title: "Homework Template"
output:
  html_document: default
  pdf_document: default
date: "2023-10-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*The following is a template .rmd RMarkdown file for you to use for your homework submission.*

*Please Knit your .rmd to a PDF format or HTML and submit that with no identifiers like your name.*

*To create a PDF, first install tinytex and load the package. Then press the Knit arrow and select "Knit to PDF".*

## QUESTION 01: Data Visualisation for Science Communication

*Create a figure using the Palmer Penguin dataset that is correct but badly communicates the data. **Do not make a boxplot**.*

*Use the following references to guide you:*

-   [*https://www.nature.com/articles/533452a*](https://www.nature.com/articles/533452a){.uri}
-   [*https://elifesciences.org/articles/16800*](https://elifesciences.org/articles/16800){.uri}

*Note: Focus on visual elements rather than writing misleading text on it.*

### a) Provide your figure here:

```{r bad figure code, echo=FALSE}

# There is no need to provide the code for your bad figure, just use echo=FALSE so the code is hidden. Make sure your figure is visible after you knit it.
#install.packages(c("ggplot2", "palmerpenguins"))

library(ggplot2)
library(palmerpenguins)

View(penguins_raw)
pengSex <- penguins_raw$Sex
pengMass <- penguins_raw$`Body Mass (g)`
pengMassSex <- ggplot(penguins_raw, aes(x=pengSex, y=pengMass)) +
  geom_bar(stat = "identity") +
  labs(y="pengMass (g)")


pengMassSex
```

### b) Write about how your design choices mislead the reader about the underlying data (200-300 words).

*Include references.*

The figure misleads the reader in the following ways:

-   Firstly a bar graph shouldn't be used to display this data , adding 'pengMass' is only somewhat informative if there are the same number of of data points in each category of Sex.

-   Another way is through the failure to specify which species of penguin are shown. In this case three different species are shown and it is likely that mass is more species dependent than sex dependent.

-   Mass is shown in grams however this is not an appropriate scale to use as data is given in standard form - this is not easily readable (though, this wouldn't be a problem if mass data wasn't addedtogether).

-   Although N/A data is useful to report -- for data shown in figures it is not needed and creates clutter. Additionally the reason for N/A data is unable to be shown in a figure and so creates more obscurity in any results trying to be shown.

-   The lack of reporting on any statistical significance also creates obscurity in the data being shown. It means the figure can only show correlation to an extent.

-   Clear axis labels should also be used - instead the axis are labelled what they are in the code used to generate the plot. Although the labels are computer and human readable, it is still misleading for humans and there is no need for the labels to be computer readable. The plot is also not titled, leaving it up to the audience to infer what is attempting to be shown

-   Overall the plot is unclear in the reporting of data and inconvenient for readers to consume as it the data that is shown is not shown clearly.

## QUESTION 2: Data Pipeline

*Write a data analysis pipeline in your .rmd RMarkdown file. You should be aiming to write a clear explanation of the steps as well as clear code.*

*Your code should include the steps practiced in the lab session:*

#### *Load the data*

#### *Appropriately clean the data*

#### *Create an Exploratory Figure (**not a boxplot**)*

#### *Save the figure*

#### ***New**: Run a statistical test*

#### ***New**: Create a Results Figure*

#### *Save the figure*

*An exploratory figure shows raw data, such as the distribution of the data. A results figure demonstrates the stats method chosen, and includes the results of the stats test.*

*Between your code, communicate clearly what you are doing and why.*

*Your text should include:*

-   *Introduction*

-   *Hypothesis*

-   *Stats Method*

-   *Results*

-   *Discussion*

-   *Conclusion*

*You will be marked on the following:*

### a) Your code for readability and functionality

### b) Your figures for communication

### c) Your text communication of your analysis

*Below is a template you can use.*

------------------------------------------------------------------------

### Introduction

```{r Data Exploration, echo=TRUE}

# Install Packages 
#install.packages(c("ggplot2", "palmerpenguins", "dplyr","ragg","tidyr"))

#Load Raw Input Data
library(palmerpenguins)

#file.create(c("figures")) #run if permissions are granted to create files - otherwise create a "figures" file.
InputData <- penguins_raw
head(InputData) 

```

The above code is designed to:

-   set up filesa  so that it can be used later in the code so figures can be added to them;

-   re-name the penguins_raw data set so that it is clearer that it is the input data

-   display the top of the input data data frame, so that variables within the data frame can be seen.

### Hypothesis

```{r Explanatory Plot, echo=TRUE}


#Clean Raw Data
library(dplyr)
library(tidyr)

CleanData <- InputData %>%
  select(all_of(c("Species","Body Mass (g)","Sex"))) %>%
  na.omit(InputData) %>%
  mutate(Species = replace(Species, Species == "Adelie Penguin (Pygoscelis adeliae)","Adelie penguin (Pygoscelis adeliae)")) %>%
  separate(col='Species', into=c('SpeciesCommon', 'SpeciesBinomial'), sep='penguin',convert = TRUE)
head(CleanData)

#Plots ExplanatoryFigure  
library(ggplot2)

ExplanatoryFigure <- ggplot(CleanData, aes(x=SpeciesCommon, y=`Body Mass (g)`, fill=Sex)) +
  geom_violin() +
  labs(x = "Species")


#Saves ExplanatoryFigure as png
library(ragg)

agg_png("figures/ExplanatoryFigure.png", 
        width = 40, 
        height = 30, 
        units = "cm", 
        res = 600, 
        scaling = 3.6)
ExplanatoryFigure
dev.off()
ExplanatoryFigure

```

I chose to look more closely on the effects of sex and species on the body mass of penguins after looking at the input data set.

The above code:

-   removes all columns that are irrelevant to my aimed analysis (omitting all data bar data on species, sex, and body mass) so that the data set can be called on without having to load irrelevant information

-   removes N/A data so that data being used for this analysis is complete

-   plots an explanatory figure so initial trends can be observed regarding the relationship between sex and species on body mass.

-   saves this explanatory plot to the figures folder previously created, so that this output can be stored and exported if needed.

### Statistical Methods

```{r Statistics, echo=TRUE}

#T-test to compare signifcance between Sex within each Species 
t_testResults <- CleanData %>%
  group_by(SpeciesBinomial) %>%
  summarise(ResultSummary = list(t.test(`Body Mass (g)` ~ Sex)))
mutate(t_testResults, p.value = c(t_testResults$ResultSummary[[1]]$p.value, t_testResults$ResultSummary[[2]]$p.value, t_testResults$ResultSummary[[3]]$p.value))

model <- lm(`Body Mass (g)` ~ Sex * SpeciesBinomial, data = CleanData)
anova_table <- anova(model)
anova_table
anova_p_value <- anova_table[["Pr(>F)"]][3]
anova_F_stat <-  anova_table[["F value"]][3] 

```

To analyse if the difference between body mass between sexes within each species is significant, and to analyse if body mass between each sex across the three species is significant, t-tests and an anova is carried out.

The t-test compares the means of body mass between sexes for each of the spices all reporting a statistically significant result as p \<0.05.

The anova compares the mean body mass across both sex and species and shows a significant difference as p \<0.01), and that sex and species likely has compounding effects on body mass as F=8.76 (3sf).

The above code:

-   carries out t-tests comparing means between sexes for each species and saves the t-test results (a list) as a data frame, t_testResults. This is done to organise the data in one place rather than having three separate lists

-   extract the p-values from the lists stored in the t_testResults data frame and adds it as a new column in the t_testResults data frame. This is so it can be called upon easier, as it is no longer nested within a list found in the data frame.

-   models the effects of sex and species on body mass and their possible compounding effects.

-   runs an analysis of variance (anova) of this model and saves this in a data frame so that data can be called open lateer

-   extracts the p-value and F-statistic from this data and saves it as separate variable so that these variables can be called upon later.

### Results & Discussion

```{r Plotting Results, echo=TRUE}

# Plots reuslts figure adding data from t-test aswell as making the explanatory figure readable for an audience. 
library(grid)

ResultsFigure <- ExplanatoryFigure +
  stat_summary(fun = mean, geom = "point", shape = 20, size = 3) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  annotate("text", x = 1, y = 6500,
           label = ifelse(t_testResults$p.value[1] < 0.05, "*", ""),
           color = "red",size = 7) +
  annotate("text", x = 2, y = 6500,
           label = ifelse(t_testResults$p.value[2] < 0.05, "*", ""),
           color = "red",size = 7) +
  annotate("text", x = 3, y = 6500,
           label = ifelse(t_testResults$p.value[3] < 0.05, "*", ""),
           color = "red",size = 7)+
  ggtitle("Violin plot of Body Mass of penguins grouped Sex and Species") +
  labs(subtitle = " Statistical significance between Sex within Species of p < 0.5 shown by * \n ANOVA showing the statistical significance across Species.") +
  theme_classic()+
  theme(plot.subtitle = element_text(
      size = 10,
      face = "italic"))
add_ANOVA <- function(figure) { 
  grid.newpage()
  grid.draw(figure)
  grid.text(label = paste("ANOVA Results"), 
          x = unit(0.93, "npc"), y = unit(0.24, "npc"), just = "center", gp = gpar(fontsize = 10))
  grid.text(label = paste("F =", round(anova_F_stat, 2)),
          x = unit(0.91, "npc"), y = unit(0.19, "npc"), just = "center", gp = gpar(fontsize = 10))
  grid.text(label = ifelse(anova_p_value < 0.01, "p < 0.01", ""),
          x = unit(0.91, "npc"), y = unit(0.15, "npc"), just = "center", gp = gpar(fontsize = 10))
}
add_ANOVA(ResultsFigure)

# Adding data from anova to Results Figure and saves figure as .png
agg_png("figures/ResultsFigure.png", 
        width = 9, height = 5, units = "cm", 
        res = 600, scaling = 0.46)
add_ANOVA(ResultsFigure)
dev.off()

```

A results figure is constructed using the explanatory figure as a skeleton, adding data from the t-tests and anova as well as changing the aesthetics of the figure to make it more readable to an audience.

The results figure shows that body mass is significantly different between all sexes within each species as the error bars displayed do not overlap. This is supported by the statistical tests reported shown by '\*' reporting p \< 0.05. Both of these features of the results figure shows a statisitcally significant difference between body mass of males and females within each species.

The anova F-statistic = 8.76 (which is significant larger than 1) tells us that the difference mean body mass between each species is greater than expected and that sex and species likely have compounding effects on body mass a p \<0.01 tells us that there is a statically significant difference between atlest one mean body mass of each sex between species.

The above code:

-   adds t-test p-value data as '\*' if the p-vlaue is \<0.05 this is done to report the statistical tests done within each species

-   adds a title and short description (as a subtitle) of the figure so that the figure is easier to understand by the audience, even with the added information fro the statistical tests

-   adds the 'classic' theme to the results figure so that it is clearer to read by the audience. The subtitled used as a description is also altered to make it smaller and italic so that the description fits on the figure and so it is clear to distinguish from the title

-   adds the anova data by creating a function. Anova data is added so that the statistical difference between of body mass between species is reported. This is added by making a function because the anova data is added as to the grid the figure is displayed on rather than the actual figure - when saving the figure as a png later the anova data will need to be added again so creating a function avoid needing to rewrite the code. The function is kept on the same .rmd file as it is only called on once and it is not too long so doesn't create to much clutter; I also didn't want to make the anova data optional to view, as it is integral to the analysis I have chose to carry out - hence I chose to keep it on the same .rmd file.

-   saves the results figure as a png to the 'figures' folder previously created so that it can be saved externally and exported if needed.

### Conclusion

Along side the t-test the anova, shows that both sex and species has an effect on body mass for the three species of penguin in the palmerpenguins data set.

------------------------------------------------------------------------

## QUESTION 3: Open Science

### a) GitHub

*Upload your RProject you created for **Question 2** and any files and subfolders used to GitHub. Do not include any identifiers such as your name. Make sure your GitHub repo is public.*

*GitHub link:*

*You will be marked on your repo organisation and readability.*

### b) Share your repo with a partner, download, and try to run their data pipeline.

*Partner's GitHub link:*

*You **must** provide this so I can verify there is no plagiarism between you and your partner.*

### c) Reflect on your experience running their code. (300-500 words)

-   *What elements of your partner's code helped you to understand their data pipeline?*

-   *Did it run? Did you need to fix anything?*

-   *What suggestions would you make for improving their code to make it more understandable or reproducible, and why?*

-   *If you needed to alter your partner's figure using their code, do you think that would be easy or difficult, and why?*

### d) Reflect on your own code based on your experience with your partner's code and their review of yours. (300-500 words)

-   *What improvements did they suggest, and do you agree?*

-   *What did you learn about writing code for other people?*
